Forecasting the 2024 U.S. Presidential Election through Poll Aggregation and Adjustments for Poll Quality

### 2024-11-04 paper/paper.qmd
fix grammar

---
title: "Forecasting the 2024 U.S. Presidential Election through Poll Aggregation and Adjustments for Poll Quality"
subtitle: "A Hybrid Model Predicts a Narrow Lead for Kamala Harris in Swing States as Election Day Approaches"
author: 
  - Colin Sihan Yang
  - Lexun Yu
  - Siddharth Gowda
thanks: "Code and data are available at: [https://github.com/yulexun/uselection](https://github.com/yulexun/uselection)."
date: today
date-format: long
abstract: "This paper develops a hybrid model to forecast the 2024 U.S. presidential election by combining poll aggregation techniques with additional variables, including pollster reliability, sample size, timing relative to the election, geographic region, and polling methodology. By integrating these factors, our model accounts for variations in poll quality and regional differences in voter sentiment, providing a more stable and accurate prediction compared to traditional poll aggregation alone. Our results show that Kamala Harris holds a slight lead over Donald Trump in most swing states, with support levels stabilizing closer to election day. This approach underscores the importance of nuanced forecasting models in election analysis, revealing how poll quality and timing can significantly influence support predictions. Ultimately, this model offers a comprehensive tool for understanding the dynamics of voter sentiment and improving the accuracy of election predictions."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

The remainder of this paper is structured as follows: @sec-data provides an overview and exploration of the data. @sec-model provides the modeling approach, including simple and multiple linear regression models and a Bayesian hierarchical model. We then present our results in @sec-result and discuss the implications, limitations, and future research directions in @sec-discussion.

@@

The dataset we obtained from @fivethirtyeight is accumulated from multiple polls and surveys. According to FiveThirtyEight, they aggregate polling data conducted by other firms and organizations that meets their methodological and ethical standards [@morris_trump_2024]. The dataset contains a list of questions in polls and surveys and their results. When a new poll is conducted, the poll is appended to the dataset. FiveThirtyEight assign the poll a poll_id, and each question is assigned a question_id.

@@

In this dataset, each row represents a polling question that records the variables of interest. Each entry allows us to explore the real-world relationships between polling factors and the support percentage (`pct`) for the candidates Kamala Harris and Donald Trump. This dataset enables an analysis of how various polling characteristics influence the reported support levels for the candidates we are focused.

@@

## Data Exploration {#sec-dataexplore}

The raw data from FiveThirtyEight contains 52 columns, all of the column headers are displayed below: 
```{r}
#| echo: false
#| warning: false
raw_data <- read_csv(here("data/01-raw_data/president_polls.csv"))

@@

These columns can be categorized into three types, response variable, numeric predictors and categorical predictors.

@@

The first 6 rows of the dataset is displayed in @tbl-cleaned-data.

@@

@fig-pollscore-samplesize illustrates a weak positive correlation between a pollster's pollscore and the sample size of their poll. The figure also shows that most polls have a sample size around 800 to 1200 participants. It is also important to note that a few polls with exceptionally large sample sizes were excluded from the graph due to their status as clear outliers.

@@

Based on @fig-polldate-canidate-pct, Trump and Harris are getting more votes in polls that are done closer to the election. This is a result of non-major candidate support rapidly decreasing. Specifically, Trump support is increasing at a faster rate than Harris. However, the polls in general show a slight lead for Harris throughout the last 100 days.

The rapid decline in third party support could be due to Robert F. Kennedy dropping out of the race (this sentences should probably be in results or discussion section).

@@

The Bayesian model is implemented in R [@citeR] using the rstanarm package as described by @rstanarm. The model is run with the following specifications:

@@

We first set up a SLR model for predicting pct based on pollscore. The model is:

@@

The primary concern lies in the evident dispersion of data points, which are widely spread and do not cluster closely around the line of perfect prediction (the dashed red line). This suggests that while pollscore may have some predictive capability, it does not adequately explain the variability in pct. The observed inconsistencies between actual and predicted values indicate that the relationship between pct and pollscore is likely not sufficiently captured by a linear model with just one predictor.

@@

In @fig-model-pct-sample_size, the distribution of the points suggests that the model might not capture much variability in pct when sample_size is the only predictor. This indicate that sample_size has limited impact on pct.

@@

Next, we predicts pct using pollscore, days teken from election, methodology, sample size, and state as predictors. The model is:

@@

The overall distribution of points suggests that the MLR model has captured a substantial portion of the variance in pct. The majority of the data points align relatively well with the red dashed line, particularly within the middle range of pct values (approximately between 40 and 60). This alignment indicates that the model performs reasonably well in this range, with predicted values correlating strongly with the actual observed outcomes.

@@

The observed data (y) follows a distinct, symmetrical distribution centered around a specific value, suggesting a well-defined peak with tapering on both sides. The replicated distributions (y_rep), shown as multiple thin lines, generally align with the shape of the observed distribution, indicating that the Bayesian model has captured the main characteristics of the data. However, the variability in the y_rep lines highlights the degree of uncertainty inherent in the model’s predictive capability.

The fact that the y_rep curves closely match the overall pattern of the actual y suggests that the model performs reasonably well in replicating the observed data. Minor discrepancies or deviations between the y and y_rep might imply areas where the model could be fine-tuned or adjusted to improve accuracy. Overall, this PPC indicates that the model provides a decent fit to the data, with some variability accounted for in the predictive samples.

@@

In addition, the residual plot @fig-bayesian-model-residual are fairly centered around zero with no major trend, suggesting that the model is not heavily biased in its predictions.

In conculsion, the visual checks from the predicted vs. actual plot and the residual plot, there is no strong evidence that the Bayesian model is overfitting. It appears to generalize well to the data it was trained on without showing signs of capturing noise or irrelevant patterns.

@@

We chose sample size, poll score, and days taken from the election as predictors for building the model based on the visual trends observed in the plots. As shown in @fig-pollsamplesize-canidate-pct, @fig-pollscore-candidate-pct and @fig-polldate-canidate-pct in @sec-data, The relationship between pct and sample size shows that larger sample sizes correspond to more stable percentages, indicating that sample size could significantly impact prediction accuracy. The plot of pct versus poll score reveals a negative trend, suggesting that poll scores are consistently influential in explaining variations in pct. Additionally, the pct versus days taken plot indicates that the proximity to the election affects the percentage, with a slight negative trend showing that time remaining has predictive value. These observed patterns imply that these variables can effectively capture key factors influencing pct and are suitable for model inclusion.

To predict the support percentage (PCT) for Harris and Trump, we employed a comprehensive modeling strategy involving Simple Linear Regression (SLR), Multiple Linear Regression (MLR), and a Bayesian hierarchical model. The SLR model provided a foundational analysis of the relationship between PCT and pollscore, highlighting its limited predictive power due to a lack of complexity. The MLR model improved on this by incorporating additional predictors such as days taken from the election, sample size, methodology, and state, enhancing its predictive accuracy and capturing interactions among variables. The Bayesian hierarchical model further refined our approach, incorporating prior knowledge and accounting for variability at group levels (e.g., methodology and state) through random intercepts. This model provided robust uncertainty quantification through credible intervals and demonstrated strong predictive performance with an R-squared value around 0.70, indicating it effectively explained data variability. By leveraging the strengths of these models, particularly the Bayesian approach’s interpretability and robustness, we achieved reliable predictions and a deeper understanding of the factors influencing support percentages.

# Result {#sec-result}
```{r}
#| label: fig-bayesian-model-CI
#| fig-cap: "Estimated Posterior Distribution for the Intercept Parameter in the Bayesian Model"
#| echo: false
#| warning: false
# Plot random effects
plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
```

@@

The estimated intercept, which represents the baseline percentage of support for Kamala Harris in the U.S. election data analyzed, is centered around 52%, with a credible interval spanning approximately from 49.5% to 53.5%. The relatively narrow width of the interval implies a certain degree of confidence in the estimate, indicating that the data used in the model provided a clear signal for the intercept's value.

@@

@fig-bayesian-model-result-CI shows that he model's predictive performance for Harris's election data appears generally robust, demonstrating a strong alignment between actual and predicted vote percentages. The majority of the data points cluster around the line, indicating reliable predictions, some variance is evident, particularly in the lower and higher actual percentage ranges. This suggests that while the model performs well on average, it may have larger uncertainty in its predictions at the extremes. @tbl-bayesian-model-result-summary also highlights that most predicted values align closely with actual values, supporting the model’s robustness in capturing the overall trend of election outcomes.

Our approach in models applies simple linear regression, multiple linear regression and Bayesian hierarchical modeling that includes both fixed and random effects, capturing variations in poll reliability. The results shows that a model accounting for poll quality, timing, and methodology can produce forecasts that are more resilient to biases found in individual polls.

One of the main takeaways from this study is the impact of poll quality on election forecasts. We weight and filtered the polls by their reliability. This approach highlights the importance of screening and adjusting for poll reliability in election forecasts, particularly in a high-stakes environment where even minor polling biases can influence public perception and campaign strategies.

## Sources of Bias In the Polls

Finally, as illustrated in Figure @fig-pollscore-candidate-pct, more reliable polls tend to show a bias favoring Trump over Harris—not necessarily indicating a guaranteed victory for Trump, but rather suggesting he may outperform expectations in many polls. It is important to note that the 538 poll score system is partly based on historical accuracy. Given Trump's overperformance in the polls during the 2016 and 2020 elections, this scoring may be overcorrecting for past inaccuracies. This potential undervaluation of Trump’s support could be strategically beneficial for the Harris campaign, providing a warning to avoid repeating previous electoral mistakes.

While this model presents a more comprehensive approach, there are limitations. First, Any systematic biases inherent in those polls can carry over into the model’s predictions because we are using existing poll data. For instance, if certain demographic groups are consistently underrepresented, our model may not fully capture their impact on election outcomes.

Second, the model only reflect current voting dynamics as it is based on historical data. This model does not account for short-term variability such as Biden's exit. This limitation is important in the final weeks before an election when small shifts in polling data can produce exaggerated effects.

On top of that, the results can be sensitive to the choice of priors. We choose normal(0, 2.5) as we believe it could stabilize the estimates by being informative enough to guide the posterior distribution without overpowering the data. However, if the chosen prior does not align well with the true underlying distribution, it may lead to biased results or overly wide credible intervals.

Each state has a limited number of polls available; while the overall analysis considered a substantial number of polls, individual states often had fewer than 20 polls, with even the largest states contributing only up to 27. Moreover, as noted in Section @sec-prediction, the analysis focused solely on closely contested races, specifitatiocally swing states. Consequently, many states were assigned to a candidate based on strong historical trends rather than direct predictions. Although this approach is reasonable, it may overlook emerging trends that could potentially shift a state's allegiance from Republican to Democrat or vice versa.

## Future Directions
To enhance the reliability and robustness of this forecasting model, future work should focus on validating its performance across multiple election cycles. We includes previous year's polling dataset in this R project. By applying the model to past elections, researchers can assess its accuracy in different political contexts and electoral dynamics. This approach would show any limitations specific to certain election conditions, such as shifts in voter demographics or the influence of emerging media platforms on public opinion. Expanding the model’s validation across multiple election cycles could help verify its robustness and adaptability, ultimately refining its accuracy and reliability for future applications in election forecasting.

# Trump Voter Prediction Model

@@

```{r}
#| label: fig-model-trumpmlr
#| fig-cap: "MLR Trump Model Accounts For A Large Amount of Variability in Voter Percentage"
#| echo: false
#| warning: false
just_trump <- cleaned_data %>% filter(candidate_name == "Donald Trump")
predictions2 = posterior_predict(bayesian_model_train_trump, newdata = just_trump)
predicted_means = colMeans(predictions2)
predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
result_summary = data.frame(
  Actual_PCT = just_trump$pct,
  Predicted_PCT = predicted_means,
  Lower_CI = predicted_intervals[1, ],
  Upper_CI = predicted_intervals[2, ]
)
# kable(head(result_summary, 10), caption = "Head of Bayesian Model Result Summary")
ggplot(result_summary, aes(x = result_summary$Actual_PCT, y = result_summary$Predicted_PCT)) +
  geom_point(color = "blue", alpha = 0.7) +
  # geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(
    title = "Actual vs Predicted Percentage For Trump MLR Model (All Trump Data)",
    x = "Actual PCT",
    y = "Predicted PCT"
  )
```

From @fig-model-trumpmlr it is clear that the model accounts for large amount of the variance in Trump's voter percentage as the data points appear close to the prediction (red line). Furthermore, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.
```{r}
#| label: fig-model-trumpmlr-res
#| fig-cap: "MLR Trump Model does not Appear to Overfit"
#| echo: false
#| warning: false
predictions2 = posterior_predict(bayesian_model_train_trump, newdata = test_data_trump)
plot(test_data_trump$pct, colMeans(predictions2), 
     xlab = "Actual PCT Values", 
     ylab = "Predicted  PCT Values", 
     main = "Predicted PCT vs Actual PCT (Test Data)")
abline(0, 1, col = "red")  # Add a 45-degree reference line
```

Based on @fig-model-trumpmlr-res, the test data predictions are also close to the actual values. This suggests that model can generalization to outside data. Similarly, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias. 
```{r}
#| label: fig-trumpmodelci
#| fig-cap: Donald Trump Expected to Recieve Approximatley 45% of the Vote
#| echo: false
#| warning: false
plot(bayesian_model_train_trump, pars = "(Intercept)", prob = 0.95)
```

@@

# Election Prediction {#sec-prediction}

Our prediction process consists of two primary components. First, we develop models for both Trump and Harris based on the variables outlined in @sec-model. This involves partitioning the dataset into training and testing subsets. Next, we further divide the testing dataset into swing states and other competitive races. We then input this test data into the respective models to generate predictions. By averaging these predictions, we can calculate the expected voter percentage for each candidate in each state. The candidate with the higher percentage is deemed the winner for that state.

@@

According to Figure @fig-currentstateofelection, Kamala Harris is predicted to be the 47th President of the United States. There are also a few states with predictions not visible in the map, we will describe those predicts them below.

@@

@@

## Brief Description of American Federal Goverment

The American federal or national government is split into three branches, executive., legislative, and judical The executive branch includes the president and the military. The legislative branch include two subgroups, the House of Representatives and the Senate. These two subgroups create laws. Every state has two Senators and one Representative per approximately 750,000 people. The judicial branch is court system.

@@

The Electoral College is the system used in the United States to elect the president and vice president. Instead of a direct popular vote, each state is allocated a certain number of electors based on its representation in Congress (the total of its Senators and Representatives). When voters cast their ballots, they are actually voting for a slate of electors pledged to a candidate. The candidate who receives a majority of electoral votes (270 out of 538) wins the presidency. This system means that winning the popular vote in a state generally results in winning all of that state's electoral votes. The only exception are the states of Maine and Nebraska, who award electoral votes by congressional district, with two additional votes given to the statewide winner.

The Electoral College results in some states being unnecessary to campaign in, as their strong historical voting pasterns towards either Democrats or Republicans make them unlikely to change, regardless of campaign efforts. Therefore, for statisticians, polling information from these states may not be that useful when trying to predict the outcome of an election. On the other hand, states that can vote either Democratic or Republican (swing states) are immensely important when predicting an election. As a result, campaigns spend hundreds of millions of dollars campaigning and understanding voters there.

```{r}
#| label: fig-currentstateofelection
#| fig-cap: 2024 U.S. Presidential Election State Forecast Map
#| echo: false
#| warning: false
light_blue <- "#ADD8E6"
dark_blue <- "#6688CC"
light_red <- "#FF7777"
dark_red <- "#B22222"
purple <- "#9370DB"
# Define your colors for each state (replace these colors with your desired ones)
state_colors_legend <- c(
  "Republican" = dark_red,      # Dark Red
  "Lean Republican" = light_red, # Light Red
  "Toss-Up" = purple,         # Purple
  "Lean Democrat" = light_blue,   # Light Blue
  "Democrat" = dark_blue,        # Dark Blue
  "Split By District" = "gray"
)

@@

The state statuses presented in this map are based on evaluations from @cnn_electoral_nodate, @fox_news_2024_2024, and @nbc_news_choose_nodate, which are generally agreed upon within the American political community. These organizations assessed historical voting patterns and recent polling data to derive their conclusions.

Notably, Nebraska and Maine are indicated in gray due to their delegates being split by district. In Nebraska, the state overall is projected to lean Republican, with the first and third districts also strongly favoring Republican candidates, while the second district leans Democratic. In Maine, the overall expectation is a Democratic leaning, consistent with its first district, though the second district leans Republican. Furthermore, Alaska and Hawaii are not in the map. Alaska is strongly favoring Republicans while Hawaii strongly favors democrats.

The seven generally agreed upon swing states are Pennsylvania, North Carolina, Georgia, Arizona, Nevada, Wisconsin, and Michigan with Texas, Florida, Nebraska District Two, Maine District 2, and Minnesota as the next closest races.

@@

# The Ideal Survey

@@

Given the monetary constraint, the team's first decision is to mostly include swing states and districts. These are Nevada, Arizona, Wisconsin, Michigan, Pennsylvania, North Carolina, and Georgia. Swing states are the primary focus of the poll because they disproportionately affect the outcome of the election based on the explanation in @sec-electoral-college. However, the poll will also include Texas, Florida, and Minnesota, Nebraska District 2, and Maine District 2, as they have the potential to one by either party but are not as likely to change as the previously mentioned swing states. Nevertheless, it is important to note that to save costs, the team will focus on polling true swing states compared to Texas, Florida and Minnesota.

@@

The team's ideal survey will be made using @qualtrics_qualtrics_nodate. It will attempt to ensure four things: no leading questions, no question order bias, no answer order bias, and the survey should identify non-engaged participants. To identify participants who are not engaged, the research team has decided to add a worthless question. This question is extremely simple and clearly has one correct answer. Therefore, if a participant selects the wrong answer, they most likely are not engaged with the survey and their responses should be removed from the final data. An example of this is question number 7.

@@

## Potential Problems with the Methodology And Polls

>>>>

---
title: "Forecasting the 2024 U.S. Presidential Election through Poll Aggregation and Adjustments for Poll Quality"
subtitle: "A Hybrid Model Predicts a Narrow Lead for Kamala Harris in Swing States as Election Day Approaches"
author: 
  - Colin Sihan Yang
  - Lexun Yu
  - Siddharth Gowda
thanks: "Code and data are available at: [https://github.com/yulexun/uselection](https://github.com/yulexun/uselection)."
date: today
date-format: long
abstract: "This paper develops a hybrid model to forecast the 2024 U.S. presidential election by combining poll aggregation techniques with additional variables, including pollster reliability, sample size, timing relative to the election, geographic region, and polling methodology. By integrating these factors, our model accounts for variations in poll quality and regional differences in voter sentiment, providing a more stable and accurate prediction compared to traditional poll aggregation alone. Our results show that Kamala Harris holds a slight lead over Donald Trump in most swing states, with support levels stabilizing closer to election day. This approach reveals how poll quality and timing can significantly influence support predictions. This model offers a comprehensive tool for understanding the dynamics of voter sentiment and improving the accuracy of election predictions. "
format: pdf
number-sections: true
bibliography: references.bib
---

@@

The remainder of this paper is structured as follows: @sec-data provides an overview of the data. @sec-model provides the modeling approach, including simple and multiple linear regression models and a Bayesian hierarchical model. We then present our results in @sec-result and discuss the implications, limitations, and future research directions in @sec-discussion.

@@

The dataset we obtained from @fivethirtyeight is accumulated from multiple polls and surveys. According to FiveThirtyEight, they aggregate polling data conducted by other firms and organizations that meets their methodological and ethical standards [@morris_trump_2024]. The dataset has a list of questions in polls and surveys and their results. When a new poll is conducted, the poll is appended to the dataset. FiveThirtyEight assign the poll a poll_id, and each question is assigned a question_id.

@@

In this dataset, each row represents a polling question that records the variables of interest. Each entry allows us to examine the real-world relationships between polling factors and the support percentage (`pct`) for the candidates Kamala Harris and Donald Trump. This dataset enables an analysis of how various polling characteristics influence the reported support levels for the candidates we are focused on.

@@

## Data Examination {#sec-dataexplore}

The raw data from FiveThirtyEight contains 52 columns, all the column headers are displayed below: 
```{r}
#| echo: false
#| warning: false
raw_data <- read_csv(here("data/01-raw_data/president_polls.csv"))

@@

These columns can be categorized into three types, response variable, numeric predictors, and categorical predictors.

@@

The first 6 rows of the dataset are displayed in @tbl-cleaned-data.

@@

@fig-pollscore-samplesize illustrates a weak positive correlation between a pollster's pollscore and the sample size of their poll. The figure also shows that most polls have a sample size of around 800 to 1200 participants. It is also important to note that a few polls with exceptionally large sample sizes were excluded from the graph due to their status as clear outliers.

@@

Based on @fig-polldate-canidate-pct, Trump and Harris are getting more votes in polls that are done closer to the election. This is a result of non-major candidate support decreasing. Specifically, Trump support is increasing at a faster rate than Harris. However, the polls in general show a slight lead for Harris throughout the last 100 days.

@@

The Bayesian model is implemented in R [@citeR] using the rstanarm package as described by @rstanarm. We run the model with the following specifications:

@@

We first set up an SLR model for predicting pct based on pollscore. The model is:

@@

The primary concern lies in the evident dispersion of data points, which are widely spread and do not cluster closely around the line of perfect prediction (the dashed red line). This suggests that while pollscore may have some predictive capability, it does not adequately explain the variability in pct. The observed inconsistencies between actual and predicted values indicate that the relationship between pct and pollscore is not sufficiently captured by a linear model with just one predictor.

@@

In @fig-model-pct-sample_size, the distribution of the points suggests that the model might not capture much variability in pct when sample_size is the only predictor. This indicates that sample_size has limited impact on pct.

@@

Next, we predict pct using pollscore, days teken from election, methodology, sample size, and state as predictors. The model is:

@@

The overall distribution of points suggests that the MLR model has captured a substantial portion of the variance in pct. Most of the data points align well with the red dashed line, particularly within the middle range of pct values (between 40 and 60). This alignment indicates that the model performs well in this range, with predicted values correlating strongly with the actual observed outcomes.

@@

The observed data (y) follows a distinct, symmetrical distribution centered around a specific value, suggesting a well-defined peak with tapering on both sides. The replicated distributions (y_rep), shown as multiple thin lines, align with the shape of the observed distribution, indicating that the Bayesian model has captured the main characteristics of the data. However, the variability in the y_rep lines highlights the degree of uncertainty inherent in the model’s predictive capability.

The fact that the y_rep curves closely match the overall pattern of the actual y suggests that the model performs well in replicating the observed data. Minor discrepancies or deviations between the y and y_rep might imply areas where the model could be fine-tuned or adjusted to improve accuracy. Overall, this PPC indicates that the model provides a decent fit to the data, with some variability accounted for in the predictive samples.

@@

In addition, the residual plot @fig-bayesian-model-residual is fairly centered around zero with no major trend, suggesting that the model is not heavily biased in its predictions.

In conclusion, from the visual checks from the predicted vs. actual plot and the residual plot, there is no robust evidence that the Bayesian model is overfitting. It appears to generalize well to the data it was trained on without showing signs of capturing noise or irrelevant patterns.

@@

We chose sample size, poll score, and days taken from the election as predictors for building the model based on the visual trends observed in the plots. As shown in @fig-pollsamplesize-canidate-pct, @fig-pollscore-candidate-pct and @fig-polldate-canidate-pct in @sec-data, the relationship between pct and sample size shows that larger sample sizes correspond to more stable percentages, indicating that sample size could significantly impact prediction accuracy. The plot of pct versus poll score shows a negative trend, suggesting that poll scores are consistently influential in explaining variations in pct. Additionally, the pct versus days taken plot indicates that the proximity to the election affects the percentage, with a slight negative trend showing that time remaining has predictive value. These observed patterns imply that these variables can effectively capture key factors influencing pct and are suitable for model inclusion.

To predict the support percentage (PCT) for Harris and Trump, we employed a modeling strategy involving Simple Linear Regression (SLR), Multiple Linear Regression (MLR), and a Bayesian hierarchical model. The SLR model provided a foundational analysis of the relationship between PCT and pollscore, highlighting its limited predictive power due to a lack of complexity. The MLR model improved on this by incorporating additional predictors such as days taken from the election, sample size, methodology, and state, enhancing its predictive accuracy and capturing interactions among variables. The Bayesian hierarchical model further refined our approach, incorporating prior knowledge and accounting for variability at group levels (e.g., methodology and state) through random intercepts. This model provided robust uncertainty quantification through credible intervals and demonstrated strong predictive performance with an R-squared value around 0.70, indicating it effectively explained data variability. By leveraging the strengths of these models, particularly the Bayesian approach’s interpretability and robustness, we achieved reliable predictions and a deeper understanding of the factors influencing support percentages.

# Result {#sec-result}

## Harris Voter Prediction Model 
```{r}
#| label: fig-bayesian-model-CI
#| fig-cap: "Estimated Posterior Distribution for the Intercept Parameter in the Bayesian Model"
#| echo: false
#| warning: false
# Plot random effects
plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
```

@@

The estimated intercept, which represents the baseline percentage of support for Kamala Harris in the U.S. election data analyzed, is centered around 52%, with a credible interval spanning from 49.5% to 53.5%. The narrow width of the interval implies a certain degree of confidence in the estimate, indicating that the data used in the model provided a clear signal for the intercept's value.

@@

@fig-bayesian-model-result-CI shows that the model's predictive performance for Harris's election data appears robust, demonstrating a strong alignment between actual and predicted vote percentages. Most of the data points cluster around the line, indicating reliable predictions, some variance is evident, particularly in the lower and higher actual percentage ranges. This suggests that while the model performs well on average, it may have larger uncertainty in its predictions at the extremes. @tbl-bayesian-model-result-summary also highlights that most predicted values align closely with actual values, supporting the model’s robustness in capturing the overall trend of election outcomes.

## Trump Voter Prediction Model

@@

```{r}
#| label: fig-model-trumpmlr
#| fig-cap: "MLR Trump Model Accounts For A Large Amount of Variability in Voter Percentage"
#| echo: false
#| warning: false
just_trump <- cleaned_data %>% filter(candidate_name == "Donald Trump")
predictions2 = posterior_predict(bayesian_model_train_trump, newdata = just_trump)
predicted_means = colMeans(predictions2)
predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
result_summary = data.frame(
  Actual_PCT = just_trump$pct,
  Predicted_PCT = predicted_means,
  Lower_CI = predicted_intervals[1, ],
  Upper_CI = predicted_intervals[2, ]
)
ggplot(result_summary, aes(x = result_summary$Actual_PCT, y = result_summary$Predicted_PCT)) +
  geom_point(alpha = 0.7) +
  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(
    title = "Actual vs Predicted Percentage For Trump MLR Model (All Trump Data)",
    x = "Actual PCT",
    y = "Predicted PCT"
  )
```

```{r}
#| label: tbl-model-trumpmlr
#| tbl-cap: "Head of Bayesian Model Result Summary"
#| echo: false
#| warning: false

kable(head(result_summary, 10))

From @fig-model-trumpmlr and @tbl-model-trumpmlr the model accounts for large amount of the variance in Trump's voter percentage as the data points appear close to the prediction (red line). Furthermore, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.

```{r}
#| label: fig-model-trumpmlr-res
#| fig-cap: "MLR Trump Model does not Appear to Overfit"
#| echo: false
#| warning: false
predictions2 = posterior_predict(bayesian_model_train_trump, newdata = test_data_trump)
plot(test_data_trump$pct, colMeans(predictions2), 
     xlab = "Actual PCT Values", 
     ylab = "Predicted  PCT Values", 
     main = "Predicted PCT vs Actual PCT (Test Data)")
abline(0, 1, col = "red")  # Add a 45-degree reference line
```

Based on @fig-model-trumpmlr-res, the test data predictions are also close to the actual values. This suggests that model can generalize outside data. Similarly, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias. 
```{r}
#| label: fig-trumpmodelci
#| fig-cap: Donald Trump Expected to Recieve Approximatley 45% of the Vote
#| echo: false
#| warning: false
plot(bayesian_model_train_trump, pars = "(Intercept)", prob = 0.95)
```

@@

## Election Prediction {#sec-prediction}

Our prediction process consists of two primary components. First, we develop models for both Trump and Harris based on the variables outlined in @sec-model. This involves partitioning the dataset into training and testing subsets. Next, we further divide the testing dataset into swing states and other competitive races. We then input this test data into the respective models to generate predictions. By averaging these predictions, we can calculate the expected voter percentage for each candidate in each state. The candidate with a higher percentage is deemed the winner for that state.

@@

According to Figure @fig-currentstateofelection, Kamala Harris is predicted to be the 47th President of the United States. There are also a few states with predictions not visible on the map, we will describe those predictions below.

@@

Our approach in models applies simple linear regression, multiple linear regression and Bayesian hierarchical modeling that includes both fixed and random effects, capturing variations in poll reliability. The results show that a model accounting for poll quality, timing, and methodology can produce forecasts that are more resilient to biases found in individual polls.

One of the main takeaways from this study is the impact of poll quality on election forecasts. We weight and filter the polls by their reliability. This approach highlights the importance of screening and adjusting for poll reliability in election forecasts, particularly in a high-stakes environment where even minor polling biases can influence public perception and campaign strategies.

## Sources of Bias in the Polls

Finally, as illustrated in Figure @fig-pollscore-candidate-pct, more reliable polls tend to show a bias favoring Trump over Harris—not necessarily indicating a guaranteed victory for Trump, but rather suggesting he may outperform expectations in many polls. It is important to note that the FiveThirtyEight poll score system is partly based on historical accuracy. Given Trump's overperformance in the polls during the 2016 and 2020 elections, this scoring may be overcorrecting for past inaccuracies. This potential undervaluation of Trump’s support could be strategically beneficial for the Harris campaign, providing a warning to avoid repeating previous electoral mistakes.

First, any systematic biases inherent in those polls can carry over into the model’s predictions because we are using existing poll data. For instance, if certain demographic groups are consistently underrepresented, our model may not fully capture their impact on election outcomes.

Second, the model only reflects current voting dynamics as it is based on historical data. This model does not account for short-term variability such as Biden's exit. This limitation is important in the final weeks before an election when small shifts in polling data can produce exaggerated effects.

On top of that, the results can be sensitive to the choice of priors. We choose Normal(0, 2.5) as we believe it could stabilize the estimates by being informative enough to guide the posterior distribution without overpowering the data. However, if the chosen prior does not align well with the true underlying distribution, it may lead to biased results or overly wide credible intervals.

Each state has a limited number of polls available; while the overall analysis considered a substantial number of polls, individual states often had fewer than 20 polls, with even the largest states contributing only up to 27. Moreover, as noted in Section @sec-prediction, the analysis focused solely on contested races, specifically swing states. Consequently, many states were assigned to a candidate based on strong historical trends rather than direct predictions. Although this approach is reasonable, it may overlook emerging trends that could potentially shift a state's allegiance from Republican to Democrat or vice versa.

## Future Directions

To enhance the reliability and robustness of this forecasting model, future work should focus on validating its performance across multiple election cycles. We include previous year's polling dataset in this R project. By applying the model to past elections, researchers can assess its accuracy in different political contexts and electoral dynamics. This approach would show any limitations specific to certain election conditions, such as shifts in voter demographics or the influence of emerging media platforms on public opinion. Expanding the model’s validation across multiple election cycles could help verify its robustness and adaptability, refining its accuracy and reliability for future applications in election forecasting.

@@

## Brief Description of American Federal Government

The American federal or national government is split into three branches, executive., legislative, and judicial the executive branch includes the president and the military. The legislative branch includes two subgroups, the House of Representatives, and the Senate. These two subgroups create laws. Every state has two Senators and one Representative per approximately 750,000 people. The judicial branch is the court system.

@@

The Electoral College is the system used in the United States to elect the president and vice president. Instead of a direct popular vote, each state is allocated a certain number of electors based on its representation in Congress (the total of its Senators and Representatives). When voters vote, they are voting for a slate of electors pledged to a candidate. The candidate who receives the most electoral votes (270 out of 538) wins the presidency. This system means that winning the popular vote in a state result in winning all that state's electoral votes. The only exception are the states of Maine and Nebraska, who award electoral votes by congressional district, with two additional votes given to the statewide winner.

The Electoral College results in some states being unnecessary to campaign in, as their strong historical voting pasterns towards either Democrats or Republicans make them unlikely to change, regardless of campaign efforts. Therefore, for statisticians, polling information from these states may not be that useful when trying to predict the outcome of an election. On the other hand, states that can vote either Democratic or Republican (swing states) are immensely important when predicting an election. As a result, campaigns spend hundreds of millions of dollars campaigning and understanding voters there. 
```{r}
#| label: fig-currentstateofelection
#| fig-cap: 2024 U.S. Presidential Election State Forecast Map
#| echo: false
#| warning: false
light_blue <- "#ADD8E6"
dark_blue <- "#6688CC"
light_red <- "#FF7777"
dark_red <- "#B22222"
purple <- "#9370DB"
# Define your colors for each state (replace these colors with your desired ones)
state_colors_legend <- c(
  "Republican" = dark_red,      # Dark Red
  "Lean Republican" = light_red, # Light Red
  "Toss-Up" = purple,         # Purple
  "Lean Democrat" = light_blue,   # Light Blue
  "Democrat" = dark_blue,        # Dark Blue
  "Split By District" = "gray"
)

@@

The state statuses presented in this map are based on evaluations from @cnn_electoral_nodate, @fox_news_2024_2024, and @nbc_news_choose_nodate, which are agreed upon within the American political community. These organizations assessed historical voting patterns and recent polling data to derive their conclusions.

Notably, Nebraska and Maine are indicated in gray due to their delegates being split by district. In Nebraska, the state overall is projected to lean Republican, with the first and third districts also strongly favoring Republican candidates, while the second district leans Democratic. In Maine, the overall expectation is a Democratic leaning, consistent with its first district, though the second district leans Republican. Furthermore, Alaska and Hawaii are not on the map. Alaska is strongly favoring Republicans while Hawaii strongly favors democrats.

The seven agreed upon swing states are Pennsylvania, North Carolina, Georgia, Arizona, Nevada, Wisconsin, and Michigan with Texas, Florida, Nebraska District Two, Maine District 2, and Minnesota as the next closest races.

@@

#The Ideal Survey

@@

Given the monetary constraint, the team's first decision is to mostly include swing states and districts. These are Nevada, Arizona, Wisconsin, Michigan, Pennsylvania, North Carolina, and Georgia. Swing states are the primary focus of the poll because they disproportionately affect the outcome of the election based on the explanation in @sec-electoral-college. However, the poll will also include Texas, Florida, Minnesota, Nebraska District 2, and Maine District 2, as they have the potential to one by either party but are not as likely to change as the previously mentioned swing states. Nevertheless, it is important to note that to save costs, the team will focus on polling true swing states compared to Texas, Florida, and Minnesota.

@@

The team's ideal survey will be made using @qualtrics_qualtrics_nodate. It will attempt to ensure four things: no leading questions, no question order bias, no answer order bias, and the survey should identify non-engaged participants. To identify participants who are not engaged, the research team has decided to add a worthless question. This question is extremely simple and clearly has one correct answer. Therefore, if a participant selects the wrong answer, they are not engaged with the survey and their responses should be removed from the final data. An example of this is question number 7.

@@

## Potential Problems with the Methodology and Polls
